---
title: "Demo - Translating C code to R code"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo - Translating C code to R code}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

```{r setup}
library(cairocore)
```

# C Code for 'Cairo Hello world' example from FAQ

This example was taken from the [CairoGraphics FAQ](https://www.cairographics.org/FAQ/)

```
#include <cairo.h>

int main (int argc, char *argv[]) {
  cairo_surface_t *surface = cairo_image_surface_create (CAIRO_FORMAT_ARGB32, 240, 80);
  cairo_t *cr = cairo_create (surface);
  
  cairo_select_font_face (cr, "serif", CAIRO_FONT_SLANT_NORMAL, CAIRO_FONT_WEIGHT_BOLD);
  cairo_set_font_size (cr, 32.0);
  cairo_set_source_rgb (cr, 0.0, 0.0, 1.0);
  cairo_move_to (cr, 10.0, 50.0);
  cairo_show_text (cr, "Hello, world");
  
  cairo_destroy (cr);
  cairo_surface_write_to_png (surface, "hello.png");
  cairo_surface_destroy (surface);
  return 0;
}
```

# Equivalent R Code for 'Cairo Hello world'

* Remove type specification for assignment to variables, as this is not valid syntax in R
* Remove trailing semi-colons.  Not essential to remove, but it looks nicer
* prefix enums (like `CAIRO_FONT_SLANT_NORMAL`) with the name of the enum they 
are contained in.  (there is a helper function called `enum_find()` to help do this)
* `cairo_destroy()` and other memory management functions do nothing in R.  They 
can still be in your code, they just don't do anything except print a warning message.

* We can save the image to a PNG file, or we can view the surface directly by 
  fetching the surface as a raster and plotting it

```{r}
library(cairocore)

# enum_find('CAIRO_FORMAT_ARGB32')
surface = cairo_image_surface_create (cairo_format_t$CAIRO_FORMAT_ARGB32, 240, 80)
cr = cairo_create (surface)

cairo_select_font_face (cr, "serif", cairo_font_slant_t$CAIRO_FONT_SLANT_NORMAL, cairo_font_weight_t$CAIRO_FONT_WEIGHT_BOLD)
cairo_set_font_size (cr, 32.0)
cairo_set_source_rgb (cr, 0.0, 0.0, 1.0)
cairo_move_to (cr, 10.0, 50.0)
cairo_show_text (cr, "Hello, world")

# cairo_surface_write_to_png (surface, "hello.png")
raster_out <- cairo_image_surface_get_raster(surface)
plot(raster_out)
```





