---
title: "Demo - Using Transforms to Flip Y-axis"
output: html_document
vignette: >
  %\VignetteIndexEntry{Demo - Using Transforms to Flip Y-axis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cairocore)
```


# Cairo surface origin is usually top left

On a default cairo surface, the origin in the top left corner with the y-axis
increasing *down* the page.


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a simple surface
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
w <- 500
h <- 500
surface <- cairo_image_surface_create (cairo_format_t$CAIRO_FORMAT_ARGB32, w, h);
cr      <- cairo_create (surface);


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Clear the screen to be green
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 0.6, 0.7, 0.6)
cairo_paint(cr)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw 2 squares
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 1, 0, 0)
cairo_rectangle(cr, 0, 0, 20, 20)
cairo_fill(cr)

cairo_set_source_rgb(cr, 0, 0, 1)
cairo_rectangle(cr, 30, 30, 40, 40)
cairo_fill(cr)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# plot the surface
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
plot(cairo_image_surface_get_raster(surface))
```



# Set transformation on surface so origin is in bottom left

By applying a transform to the canvas we can set the origin to be the bottom left, 
with the *y* axis increasting up the page.


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a simple surface
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
w <- 500
h <- 500
surface <- cairo_image_surface_create (cairo_format_t$CAIRO_FORMAT_ARGB32, w, h);
cr      <- cairo_create (surface);


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Flip the y axis
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
mat <- cairo_matrix_t(1, 0, 0, -1, 0, h)
cairo_set_matrix(cr, mat)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Clear the screen to be green
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 0.6, 0.7, 0.6)
cairo_paint(cr)



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw 2 squares
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 1, 0, 0)
cairo_rectangle(cr, 0, 0, 20, 20)
cairo_fill(cr)

cairo_set_source_rgb(cr, 0, 0, 1)
cairo_rectangle(cr, 30, 30, 40, 40)
cairo_fill(cr)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# plot the surface
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
plot(cairo_image_surface_get_raster(surface))
```


# Flipping Y axis applies to many things, but not the actual strokes of a font

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a simple surface
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
w <- 500
h <- 500
surface <- cairo_image_surface_create (cairo_format_t$CAIRO_FORMAT_ARGB32, w, h);
cr      <- cairo_create (surface);


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Flip the y axis
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
mat <- cairo_matrix_t(1, 0, 0, -1, 0, h)
cairo_set_matrix(cr, mat)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Clear the screen to be green
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 0.6, 0.7, 0.6)
cairo_paint(cr)



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw 2 squares
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 1, 0, 0)
cairo_rectangle(cr, 0, 0, 20, 20)
cairo_fill(cr)

cairo_set_source_rgb(cr, 0, 0, 1)
cairo_rectangle(cr, 30, 30, 40, 40)
cairo_fill(cr)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Set up some text
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_select_font_face (cr, "serif",
                        cairo_font_slant_t$CAIRO_FONT_SLANT_NORMAL,
                        cairo_font_weight_t$CAIRO_FONT_WEIGHT_BOLD);

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Invert the text to match origin at lower-left
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
font_size <- 70
cairo_set_font_size(cr, font_size)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw the text
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
text_message <- "Hello #RStats"

cairo_set_source_rgb (cr, 0.0, 0.0, 1.0);
cairo_move_to (cr, 20, 200);
cairo_show_text (cr, text_message)


plot(cairo_image_surface_get_raster(surface))

```



# Flipping Y for canvas and text

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a simple surface
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
w <- 500
h <- 500
surface <- cairo_image_surface_create (cairo_format_t$CAIRO_FORMAT_ARGB32, w, h);
cr      <- cairo_create (surface);


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Flip the y axis
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
mat <- cairo_matrix_t(1, 0, 0, -1, 0, h)
cairo_set_matrix(cr, mat)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Clear the screen to be green
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 0.6, 0.7, 0.6)
cairo_paint(cr)



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw 2 squares
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_set_source_rgb(cr, 1, 0, 0)
cairo_rectangle(cr, 0, 0, 20, 20)
cairo_fill(cr)

cairo_set_source_rgb(cr, 0, 0, 1)
cairo_rectangle(cr, 30, 30, 40, 40)
cairo_fill(cr)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Set up some text
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_select_font_face (cr, "serif",
                        cairo_font_slant_t$CAIRO_FONT_SLANT_NORMAL,
                        cairo_font_weight_t$CAIRO_FONT_WEIGHT_BOLD);

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Invert the text to match origin at lower-left
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
font_size <- 70
tmat <- cairo_matrix_t(font_size, 0, 0, -font_size, 0, 0)
cairo_set_font_matrix(cr, tmat)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw the text
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
text_message <- "Hello #RStats"

cairo_set_source_rgb (cr, 0.0, 0.0, 1.0);
cairo_move_to (cr, 20, 200);
cairo_show_text (cr, text_message)


plot(cairo_image_surface_get_raster(surface))

```

