---
title: "Demo - Vectorised Functions"
output: html_document
vignette: >
  %\VignetteIndexEntry{Demo - Vectorised Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

```{r setup}
library(cairocore)
```


# C vs R

### C is fast

* C is fast
* For loops are fast in C
* Calling functions in C is fast


### R is slower

* R is slower than C 
* For loops can be slow in R
* Calling functions in R can be slow


# Vectorised calls for avoiding for loops in R

`cairocore` includes some functions for performing multiple operations with 
a single call from R - looping happens within C, and not within R.

The current vectorised functions are:

* `cairo_arc_vec()` - draw lots of circles. A vectorised version of `cairo_arc()`
* `cairo_rectangle_vec()` - draw lots of rectangles. A vectorised version of `cairo_rectangle()`
* `cairo_segment_vec()` - draw lots of line segments. A vectorised version of `cairo_line_to()`
* `cairo_polygon_vec()` - draw lots of polygons.  No equivalent in CairoGraphics


If only drawing a few objects then the vectorised calls will not save you any 
significant time.


# Drawing 20000 rectangles at 30 frames-per-second

The following code draws 20,000 rectangles on a 1000x1000 canvas with a single call. 
This takes around 30ms - which translates to around 30 frames-per-second.

Drawing the rectangles one-at-a-time in a for loop with calls to `cairo_rectangle()`
is about about half the speed.  This speed difference probably isn't that critical
for the majority of applications, but for realtime rendering with interactivity, 
R will need every speed boost it can get.


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Set up surface. Set to a white background. Turn off antialiasing
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
size    <- 1000
surface <- cairo_image_surface_create (cairo_format_t$CAIRO_FORMAT_ARGB32, size, size);
cr      <- cairo_create (surface);
cairo_set_source_rgb(cr, 1, 1, 1)
cairo_paint (cr);

cairo_set_antialias(cr, cairo_antialias_t$CAIRO_ANTIALIAS_NONE)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Generate coords and colours for 20000 rectangles
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N <- 20000

x      <- runif(N, 1, size)
y      <- runif(N, 1, size)
width  <- runif(N, 1, size/10)
height <- runif(N, 1, size/10)
r      <- runif(N)
g      <- runif(N)
b      <- runif(N)
a      <- runif(N)
sr     <- runif(N)
sg     <- runif(N)
sb     <- runif(N)
sa     <- runif(N)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw 20000 rectangles in a single call
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
cairo_rectangle_vec(cr, x, y, width, height, r, g, b, a = 1, sr, sg, sb, sa=0)

raster_out <- cairo_image_surface_get_raster(surface)
plot(raster_out, interpolate = FALSE)
```






