---
title: "Technical - C pointers in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Technical - C pointers in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(cairocore)
```



# Pointers in C

Pointers in C are memory addresses which *point* to the location of data.

This data is created as the program runs, and it is usually the users 
responsibility to release this memory.

Issues will arise if more than 1 part of the program assumes it has responsibility 
for freeing that memory.

The following C code creates a pointer to a `cairo_rectangle_t` struct.


```
cairo_rectangle_t *my_rect = (cairo_rectangle_t *)calloc(1, sizeof(cairo_rectangle_t));
```




# Pointers in R

In R, pointers which have been allocated within C are represented as *External Pointers*.

These are opaque objects which point to a memory location, but don't show any interior
details when viewed in R.

The following R code in `cairocore` creates a pointer to a rectangle object


```{r}
my_rect <- cairo_rectangle_t(x = 10, y = 10, width = 100, height = 200)

my_rect
typeof(my_rect)
class(my_rect)
```


This external pointer can then be used when calling other functions within 
`cairocore` which expect a `cairo_rectangle_t *` as an argument.


# Freeing the memory associated with pointers

In C, the programmer is tasked with manually releasing any memory that was allocated
and assigned to a pointer - usually using the function `free()`.

However, as R is a *garbage collected* language, there's no concept which is equivalent 
to calling `free()` as would be done in C.

So there is a mismatch between how R and C want to deal with releasing memory when they're 
finished using it:  C wants the programmer to manually call a function to release the memory,
where R just lets the object fall out of scope and be garbage collected.

The approach taken to resolve this mismatch in `cairocore` is:

1. Pointers which are returned to R have *finalizer* functions attached.  Thus, when they
   fall out of scope and are removed by the garbage collector, the finalizer function
   will be called by the garbage collector to release the memory.
2. All manual methods from the original Cairo C library which are related to releasing the memory 
   are not available from R.  
   
   
These two steps ensure that R users cannot free the memory behind a pointer before the
R runtime has finished with it.
  













