---
title: "Technical - C enums in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Technical - C enums in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)

library(cairocore)
```


# Enums in C

`enums` in C are a datastructure defining a mapping from human-readable variable 
names to integer values.  For example, the
`cairo_format_t` enum in the Cairo library defines the valid colour formats 
that it knows how to deal with.

Instead of having to remember that `2` stands for `CAIRO_FORMAT_A8` when writing C code,
the programmer would just use `CAIRO_FORMAT_A8` and the compiler would automatically 
replace this with `2` everywhere it appears.


```
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
// The definition of 'cairo_format_t' in the C header 'cairo.h'
//~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
typedef enum _cairo_format {
    CAIRO_FORMAT_INVALID   = -1,
    CAIRO_FORMAT_ARGB32    = 0,
    CAIRO_FORMAT_RGB24     = 1,
    CAIRO_FORMAT_A8        = 2,
    CAIRO_FORMAT_A1        = 3,
    CAIRO_FORMAT_RGB16_565 = 4,
    CAIRO_FORMAT_RGB30     = 5,
    CAIRO_FORMAT_RGB96F    = 6,
    CAIRO_FORMAT_RGBA128F  = 7
} cairo_format_t;
```


# Enums in R

R doesn't have the concept of an enum, so in `cairocore` they are implemented 
as named lists of integers.

Each enum has documentation accessible via the standard help mechanism e.g. `?cairo_format_t`


```{r}
cairo_format_t
```


So to access an enum value by its name, you would write:

```{r}
cairo_format_t$CAIRO_FORMAT_A8
```


To look-up an enum name by its value there is a helper function `enum_lookup()` - 

```{r}
enum_lookup(cairo_format_t, 2)
```





# Compatability issues

One downside of implementing enums like this in R is that there is little in the
way of checking of values.


In C, if you tried to use `CAIRO_FORMAT_A_DOES_NOT_EXIST` then the compiler would
throw an error.

In R, using `cairo_format_t$CAIRO_FORMAT_A_DOES_NOT_EXIST` will return a `NULL`, 
which would then mean a likely error somewhere later in the program.


# Possibilities

To avoid the use of undefined enums, a new datatype/class could be defined which 
was stricter about raising an error if a supplied enum name didn't exist.













