---
title: "Technical - C structs in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Technical - C structs in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)

library(cairocore)
```


# Structs in C

`structs` in C are a data structure for combining items which may be of different types.

In Cairo, a common struct contains the data defining a single rectangle.


```
typedef struct _cairo_rectangle {
    double x, 
    double y, 
    double width, 
    double height;
} cairo_rectangle_t;
```

Elements of a struct may be accessed by name using `->` or `.` depending on whether it is
a pointer or not

```
void some_function(cairo_rectangle_t *my_rect) {
  printf("My rectangle width is %f\n", my_rect->width);
}
```


# Structs in R

`structs` in R are implemented as external pointer objects:

* they're *opaque* in the sense that you can't see into them.
* they're *pointer objects* in the sense that they are what R refers to as *External Pointers*, and 
  merely contain the address of the data in memory.

In this next example,  a rectangle struct is created in R.  Note: 

* The function for the creation of a struct is the same name as the struct itself
* The class of the returned object is the same as the struct
* The object is an external pointer

```{r}
my_rect <- cairo_rectangle_t(x = 10, y = 10, width = 100, height = 200)

my_rect
typeof(my_rect)
class(my_rect)
```


# Struct helpers in R - `as.list()`

For some structs, the `as.list()` method has been implemented on the class to 
convert the external pointer to a regular R list.

Note: This is not currently a link between the external pointer data and the list representation.  Changing 
one of the objects doesn't not change the other one.

```{r}
as.list(my_rect)
```











